---
title: "Gradient forest"
author: "Matt G. DeSaix"
date: "5/17/2022"
output: pdf_document
---

Final gradient forest stuff. The paramters *mtry* and *ntree* have been chosen based on the top importance of the tuning parameter test. Went with the parameters that provided the highest importance to the top variable, *PAS*.

mtry: 5
ntree: 100


```{r, message = F, warning = F}
library(tidyverse)
library(gradientForest)
library(terra)
library(sf)
library(ggspatial)
library(raster)
```

### combining gradient forest objects

```{r}
cand.gf <- readRDS("./tuning-params/data/bcrf.rda_lfmm_cand500.env.1991.ntree.100.mtry.5.rds")
rand.gf <- readRDS("./tuning-params/cand-random/1991_2020/bcrf.rda_lfmm_cand500.env.1991.ntree.100.mtry.5.random.8.rds")
full.gf <- readRDS("~/Downloads/bcrf.400k.env.1991.ntree.100.mtry.5.rds")

```

# Creating cumulative importance data frames

### Candidate Snps

```{r}
preds <- names(cand.gf$overall.imp)
cand.preds.list <- list()
for(i in preds){
  CU <- cumimp(cand.gf, i)
  cand.preds.list[[i]] <- data.frame(x = CU$x,
                                     y = CU$y,
                                     pred = i,
                                     snps = "cand")
}
cand.preds.df <- do.call(rbind, cand.preds.list)
```

### Random SnPs

```{r}
rand.files <- list.files("./tuning-params/cand-random/1991_2020/",
                         full.names = T)
rand.preds.list <- list()
for(i in preds){
  x.list <- list()
  for(j in 1:length(rand.files)){
    rand.gf <- readRDS(rand.files[j])
    CU <- cumimp(rand.gf, i)
    isub <- seq(1, length(CU$x), len=pmin(500,length(CU$x)))
    x.list[[j]] <- data.frame(x = as.character(CU$x[isub]), y = CU$y[isub])
    
  }
  tmp.df <- purrr::reduce(x.list, dplyr::left_join, by = "x") %>%
    drop_na()
  y <- apply(tmp.df[,2:ncol(tmp.df)], 1, mean)
  rand.preds.list[[i]] <- data.frame(x = tmp.df$x,
                                     y = y,
                                     pred = i,
                                     snps = "rand")
}
rand.preds.df <- do.call(rbind, rand.preds.list)
```

### All snps

```{r}
full.preds.list <- list()
for(i in preds){
  CU <- cumimp(full.gf, i)
  isub <- seq(1, length(CU$x), len=pmin(500,length(CU$x)))
  full.preds.list[[i]] <- data.frame(x = CU$x[isub],
                                     y = CU$y[isub],
                                     pred = i,
                                     snps = "full")
}
full.preds.df <- do.call(rbind, full.preds.list)
```


## Combine cumulative importance dfs and plot

```{r}
all.snps.df <- rbind(cand.preds.df, rand.preds.df, full.preds.df) %>%
  mutate(x = as.numeric(x))
three.colors <- c("#7fc97f", "#beaed4", "#fdc086")
y.max <- max(all.snps.df$y)

p.imp <- all.snps.df %>%
  filter(pred == "MWMT") %>%
  ggplot() +
  geom_step(aes(x=x,y=y,color=snps), size = 1) +
  scale_color_manual(values = three.colors) +
  theme_bw() +
  ylab("Cumulative Importance") +
  xlab("MWMT") +
  ylim(c(0,y.max))
# ggsave(plot = p.imp, filename = "./plots/turnover-functions/combined-snp-sets/mwmt-turnover-snps.png", width = 6, height = 4)
p.imp
```



```{r}
plot(full.gf, plot.type = "C", show.overall = T, 
     show.species = F, imp.vars = c("PAS", "MWMT", "SHM", "Elev", "MEM1"), common.scale = T)
```

```{r}
plot(full.gf, plot.type = "O")
```

### heat plot

```{r}
rand.files <- list.files("./tuning-params/cand-random/1991_2020/",
                         full.names = T)
rand.preds.list <- list()

for(i in 1:length(rand.files)){
  rand.preds.list[[i]] <- importance(readRDS(rand.files[i]), "Weighted")
}
rand.imp.mean <- do.call(rbind, rand.preds.list) %>%
  apply(., 2, mean)

weighted.r2s <- rbind(importance(full.gf, "Weighted"),
                      importance(cand.gf, "Weighted"),
                      rand.imp.mean) %>%
  as.data.frame() %>%
  add_column(Group = c("All", "Candidate", "Random")) %>%
  pivot_longer(cols = PAS:Elev, names_to = "Predictor", values_to = "Importance")

all.var.heat <- all.var.imp %>%
  filter(ntree == 100 & mtry == 5 & Year == "1991") %>%
  mutate(Predictor = factor(Predictor, levels = c("MEM1", "Elev", "SHM", "MWMT", "PAS")),
         Group = recode(Group, AllSNPs_1991 = "All",
                        CandSNPs_1991 = "Candidate",
                        CandSNPs_Random_1991 = "Random")) %>%
  rename(`SNP sets` = Group) %>%
ggplot(aes(`SNP sets`, Predictor)) +
  geom_tile(aes(fill = Importance)) +
  scale_fill_distiller(palette = "BuGn", direction = 1) +
  theme_bw()

# ggsave(filename = "./tuning-params/plots/var.imp.all.heat.bugn.png",
#        plot = all.var.heat, width = 6, height = 4)
all.var.heat
```



