---
title: "Gradient forest"
author: "Matt G. DeSaix"
date: "5/17/2022"
output: pdf_document
---

Final gradient forest stuff. The paramters *mtry* and *ntree* have been chosen based on the top importance of the tuning parameter test. Went with the parameters that provided the highest importance to the top variable, *PAS*.

mtry: 5
ntree: 100


```{r}
library(tidyverse)
library(gradientForest)
library(terra)
library(sf)
library(ggspatial)
library(raster)
```


```{r}
bcrf.gf <- readRDS("./tuning-params/data/bcrf.rda_lfmm_cand500.env.1991.ntree.100.mtry.5.rds")

sample.sites <- read_csv("../../data/ROFI-meta-tidy.csv") %>%
  filter(Alpha_Code == "BCRF") %>%
  group_by(Group_ID) %>%
  summarize(Lat = mean(Lat),
            Lon = mean(Long),
            .groups = "drop") %>%
  filter(Group_ID != "EBMO")
```

```{r}
plot(bcrf.gf)
```

### Preciptation as snow (PAS)

```{r}
plot(bcrf.gf, plot.type = "C", imp.vars = c("PAS"), 
     show.overall = F, plot.args = list(ylim = c(0,1)))
```

```{r}
plot(bcrf.gf, plot.type = "C", show.overall = T, 
     show.species = F,
     imp.vars = "PAS")
```




### Mean warmest temp

```{r}
plot(bcrf.gf, plot.type = "C", imp.vars = c("MWMT"), 
     show.overall = F, plot.args = list(ylim = c(0,1)))
```

```{r}
plot(bcrf.gf, plot.type = "C", show.overall = T, 
     show.species = F,
     imp.vars = "MWMT")
```

## Gradient forest map

```{r, echo = F}
breeding.shp <- st_read(dsn = "~/Desktop/biomod2/output/final/top6/shapefile") %>%
  vect()

# env.1961 <- c(terra::rast("../data/adaptwest_1961-1990.tif")[[c("MWMT", "PAS", "SHM")]], terra::rast("../data/adaptwest_elevation.tif")) %>%
#   terra::mask(breeding.shp)
env.1991 <- c(terra::rast("../data/adaptwest_1991-2020_res0.008.tif")[[c("MWMT", "PAS", "SHM")]], terra::rast("../data/adaptwest_elevation.tif")) %>%
  crop(breeding.shp) %>%
  mask(breeding.shp)
names(env.1991) <- c("MWMT", "PAS", "SHM", "Elev")
```

```{r}
# randompoints.bioclim <- spatSample(env.1991, size = 10000, 
#                                     method = "random",
#                                       na.rm = T, xy = T, as.df = T)
randompoints.bioclim <- as.data.frame(env.1991, xy = T)

# select everything except lat/lon
predictors_uncor <- colnames(randompoints.bioclim)[3:ncol(randompoints.bioclim)]

current_grid <- cbind(randompoints.bioclim[,c("x", "y")],
                      predict(bcrf.gf, randompoints.bioclim[,predictors_uncor] %>%
                                              as.data.frame()))
PCs <- prcomp(current_grid[,c(3:6)]) #run your PCs on your top four variables
# PCs <- prcomp(current_grid[,c(3:8)])
# set up a colour palette for the mapping
a1 <- PCs$x[,1]
a2 <- PCs$x[,2]
a3 <- PCs$x[,3]
r <- a1+a2
b <- a3+a2-a1
g <- -a2
r <- (r-min(r)) / (max(r)-min(r)) * 255
g <- (g-min(g)) / (max(g)-min(g)) * 255
b <- (b-min(b)) / (max(b)-min(b)) * 255
cols <- rgb(r,g,b,max=255)
cols2 <- col2rgb(cols)
cols3 <- t(cols2)
gradients <- cbind(current_grid[c("x","y")],cols3)


##Biplot specifics
nvs <- dim(PCs$rotation)[1]
# vec <- c("bio19", "bio3", "bio15", "ffdays", "bio11")
# vec <- c("BIO18", "BIO19")
# I added the code below
vec <- rownames(PCs$rotation)
lv <- length(vec)
vind <- rownames(PCs$rotation) %in% vec
scal <-100
xrng <- range(PCs$x[, 1], PCs$rotation[, 1]/scal) *1
yrng <- range(PCs$x[, 2], PCs$rotation[, 2]/scal) *1


```


```{r}
biplot.site <- predict(bcrf.gf)
PCsites <- predict(PCs, biplot.site) %>%
  as.data.frame() %>%
  add_column(Group_ID = sample.sites$Group_ID)

pt.cols <- rgb(r, g, b, maxColorValue = 255)
pcs.df <- as.data.frame(PCs$x) %>%
  mutate(group = paste0("Group_", PC1, "_", PC2))

line.scale <- 10
lines.env <- as.data.frame(PCs$rotation)/line.scale

p.biplot <- pcs.df %>%
  ggplot() +
  geom_point(aes(x = PC1, y = PC2, 
                 color = factor(group, levels = group)), 
             show.legend = F,
             size = 0.5, shape = 15) +
  scale_color_manual(values = pt.cols) +
  geom_point(data = PCsites, aes(x = PC1, y=PC2), 
             shape = 24, size = 6, fill = "grey", stroke = 2) +
  geom_segment(aes(x = 0, xend = PC1, y = 0, yend = PC2),
               size = 2,
               data = lines.env,
               arrow = grid::arrow(length = unit(0.2, "cm")),
               inherit.aes = F) +
  theme_bw() +
  coord_equal() +
  theme(
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "transparent",colour = NA),
        plot.background = element_rect(fill = "transparent",colour = NA),
        axis.ticks = element_blank(),
        axis.text = element_blank()
        )

# ggsave(plot = p.biplot, filename = "./plots/biplot.png", width = 8, height = 6)
p.biplot
```





```{r}
plot((PCs$x[, c(1,2)]), xlim = xrng, ylim = yrng, pch = ".", cex = 4, col = rgb(r, g, b, max = 255), asp = 1)
points(PCs$rotation[!vind, c(1,2)]/scal, pch = "+")
arrows(rep(0, lv), rep(0, lv), 
       PCs$rotation[vec,1]/scal*10, 
       PCs$rotation[vec, 2]/scal*10, length = 0.1)
jit <- 0.02
text(PCs$rotation[vec, 1]/scal + jit * sign(PCs$rotation[vec,1]), PCs$rotation[vec, 2]/scal + jit * sign(PCs$rotation[vec,2]), labels = vec)
```



### Pretty map

```{r}
library(raster)
elev <- raster("./data/bcrf-elevation-agg-270m.tif") %>%
  raster::crop(extent(-108.5, -104.8, 36.4, 42))
crs(elev) <- "+init=epsg:4326"
dsm.hill <- hillShade(terrain(elev,opt='slope'), 
                      terrain(elev,opt='aspect'),
                      angle = 40, direction = 270) %>%
  rasterToPoints() %>%
  data.frame()
gg.elev <- elev %>%
  rasterToPoints() %>%
  data.frame()
colnames(gg.elev) <- c("x","y", "Elevation")

states.shp <- st_read("../../data/GIS/ne_shapefiles/ne_10m_admin_1_states_provinces_lines.shp") %>%
  st_as_sfc()
breeding.shp <- st_read(dsn = "~/Desktop/biomod2/output/final/top6/shapefile") %>%
  st_as_sfc()
```

```{r}
bcrf.map <- as.data.frame(gradients)

gg <- ggplot() +
  geom_raster(data = gg.elev, aes(x,y,fill = Elevation), alpha = 0.9) +
  geom_raster(data = dsm.hill, aes(x,y,alpha = 1-layer),fill = "gray50") +
  scale_alpha(guide='none',range = c(0,1.00)) +
  # scale_fill_gradientn(name = "Elevation (meters)", colors = terrain.colors(100)) +
  scale_fill_gradientn(name = "Elevation (m)",
                       colors = gray.colors(n = 100,start = 0.01, end = 0.8, rev = T)) +
  geom_tile(data = data.frame(bcrf.map), 
            aes(x,y), fill = rgb(r,g,b, max = 255),
            height = 0.0223, width = 0.0223,
            show.legend = F,
            alpha = 0.75) +
   geom_sf(data = breeding.shp, fill = "transparent", color = "black", size = 0.2) +
  geom_sf(data = states.shp, fill = "transparent", color = "grey90") +
 geom_point(data = sample.sites, 
            aes(Lon,Lat), size = 3, alpha = 0.8, 
            fill = "grey", color = "black", stroke = 2, shape = 24) +
  # scale_shape_manual(values = 24, labels = "", name = "Sampling site") +
  theme_bw() +
  xlab("Longitude") +
  ylab("Latitude") +
   theme(panel.grid = element_blank(),
         axis.line = element_blank(),
         panel.background = element_blank(),
         axis.title = element_text(size = 14),
         axis.text = element_text(size = 12)) +
  xlim(c(-108.5, -104.8)) +
  ylim(c(36.4,42)) +
  coord_sf()
gg
```


### Beta map

```{r}
bcrf.map <- as.data.frame(gradients)

gg2 <- ggplot() +
  geom_raster(data = gg.elev, aes(x = x, y = y, fill = Elevation)) +
  # geom_raster(data = dsm.hill, aes(x = x, y = y, alpha = 1 - layer), fill = "gray50") +
  scale_fill_gradientn(name = "Elevation (m)",
                       colors = gray.colors(n = 256, rev = T, gamma = 1)) +
  geom_tile(data = data.frame(bcrf.map), 
            aes(x,y), fill = rgb(r,g,b, max = 255),
            height = 0.0223, width = 0.0223,
            show.legend = F,
            alpha = 0.6) +
   geom_sf(data = breeding.shp, fill = "transparent", color = "black", size = 0.2) +
  geom_sf(data = states.shp, fill = "transparent", color = "grey10") +
 geom_point(data = sample.sites, 
            aes(Lon,Lat), size = 2, alpha = 0.8, 
            fill = "grey", color = "black", stroke = 1, shape = 24) +
  # scale_shape_manual(values = 24, labels = "", name = "Sampling site") +
  theme_bw() +
  xlab("Longitude") +
  ylab("Latitude") +
   theme(panel.grid = element_blank(),
         axis.line = element_blank(),
         panel.background = element_blank(),
         axis.title = element_text(size = 14),
         axis.text = element_text(size = 12),
         legend.position = "none") +
  xlim(c(-108.5, -104.8)) +
  ylim(c(36.4,42)) +
  coord_sf()
# ggsave(plot = gg2, filename = "./plots/gf-map/gradient-forest-NOshade_gamma_1.png")
gg2
```


### customizing cumulative importance plots

```{r}
cand.gf <- readRDS("./tuning-params/data/bcrf.rda_lfmm_cand500.env.1991.ntree.100.mtry.5.rds")
rand.gf <- readRDS("./tuning-params/cand-random/1991_2020/bcrf.rda_lfmm_cand500.env.1991.ntree.100.mtry.5.random.8.rds")
full.gf <- readRDS("~/Downloads/bcrf.400k.env.1991.ntree.100.mtry.5.rds")

```


standard plot

```{r}
plot(cand.gf, plot.type = "C", show.overall = T, 
     show.species = F, imp.vars = c("PAS", "MWMT", "SHM", "Elev", "MEM1"), common.scale = T)
```

```{r}
cand.pas.dens.df <- data.frame(x = cand.gf$dens$PAS$x,
                               y = cand.gf$dens$PAS$y)

ggplot(cand.pas.dens.df, aes(x = x, y = y)) +
  geom_line()
```

